\thornado-hydro: Generalizing discontinuous galerkin methods for a nuclear equation of state for supernova hydrodynamics

### 2019-03-04 paper/ms.tex
Some Updates, Added notebook to create FORTRAN expressions.

For the sake of efficiency we let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
spectral-type nodal collocation DG approximation \citep{bassi:2013}.
Inserting Eq.~\eqref{eq:conservedNodalExpansion} into Eq.~\eqref{eq:semidiscreteDG},
letting $\phi(x)=\ell_{k}(x)$, using the quadratures defined above,
we obtain\footnote{Are these summations?}
\begin{align}
  \pd{}{t}\int_{K}\vect{U}_{h}\,\phi\,dx
  &\approx w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
  \label{eq:timeDerivativeTerm}
\end{align}
for the time derivative piece.
Similarly, the last term on the left-hand side of Eq.~\eqref{eq:semidiscreteDG}
becomes
\beq
  \int_{K}\vect{F}(\vect{U}_{h})\,\pderiv{\phi}{x}\,dx
  \approx \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
  \label{eq:volumeTerm}
\eeq
Now we may combine Eqs.~\eqref{eq:timeDerivativeTerm}-\eqref{eq:volumeTerm} in
Eq.~\eqref{eq:semidiscreteDG} resulting in the semi-discrete form
\beq
\begin{split}
  \deriv{\vect{U}_{k}}{t}
  & =-\f{1}{w_{k}\Delta x}
  \Big\{
  \Big[\,
    \widehat{\vect{F}}\ell_{k}\big|_{x_{H}}
     -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
  \,\Big] \\
  & -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
  \Big\}.
  \label{eq:semidiscreteDiscretized}
\end{split}
\eeq
We now have a system of ordinary
differential equations (ODEs), which may be evolved in time with an ODE solver.
In Section~\ref{sec:tests} we use the third-order strong
stability-preserving Runge-Kutta (SSP-RK3) method \citep{shu:1988}.

@@

@@

\section{Derivation of Jacobian and Eigenvalues for the 1D Nuclear Case}

We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
where $m = \rho u$ and $D = \rho y_{e}$. The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.

The pressure differential is given by

\begin{equation}
  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD
\end{equation}

The internal energy differential can be written as

\begin{equation}
\begin{split}
	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE
\end{split}
\end{equation}

where $p_{x} = \pderiv{p}{x}$

The elements of the Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ are given by:

\begin{equation}
  \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
  & = v(2-p_{\epsilon}\tau)
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
  & = \tau p_{\epsilon}
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
  & = H - p_{\epsilon}v^{2}\tau
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
\end{split}
\end{equation}

>>>>

For the sake of efficiency we let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
spectral-type nodal collocation DG approximation \citep{bassi:2013}.
Inserting Eq.~\eqref{eq:conservedNodalExpansion} into Eq.~\eqref{eq:semidiscreteDG},
letting $\phi(x)=\ell_{k}(x)$, using the quadratures defined above,
we obtain\footnote{Are these summations?}
\begin{align}
  \pd{}{t}\int_{K}\vect{U}_{h}\,\phi\,dx
  &\approx w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
  \label{eq:timeDerivativeTerm}
\end{align}
for the time derivative piece.
Similarly, the last term on the left-hand side of Eq.~\eqref{eq:semidiscreteDG}
becomes
\beq
  \int_{K}\vect{F}(\vect{U}_{h})\,\pderiv{\phi}{x}\,dx
  \approx \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
  \label{eq:volumeTerm}
\eeq
Now we may combine Eqs.~\eqref{eq:timeDerivativeTerm}-\eqref{eq:volumeTerm} in
Eq.~\eqref{eq:semidiscreteDG} resulting in the semi-discrete form
\beq
\begin{split}
  \deriv{\vect{U}_{k}}{t}
  & =-\f{1}{w_{k}\Delta x}
  \Big\{
  \Big[\,
    \widehat{\vect{F}}\ell_{k}\big|_{x_{H}}
     -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
  \,\Big] \\
  & -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
  \Big\}.
  \label{eq:semidiscreteDiscretized}
\end{split}
\eeq
We now have a system of ordinary
differential equations (ODEs), which may be evolved in time with an ODE solver.
In Section~\ref{sec:tests} we use the third-order strong
stability-preserving Runge-Kutta (SSP-RK3) method\footnote{Change later if necessary} \citep{shu:1988}.

@@

\section{Derivation of Jacobian and Eigensystem for the 1D Nuclear Case}

We assume that pressure $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$ is the specific volume.
Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
where $m = \rho u$ and $D = \rho y_{e}$.
The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.

The pressure differential is given by
\begin{equation}
  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD.
\end{equation}
The internal energy differential can be written as

\begin{equation}
\begin{split}
	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE.
\end{split}
\end{equation}

where $p_{x} := \pderiv{p}{x}$

The flux Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ may then be shown to be:

% \begin{equation}
%   \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
% \end{equation}
%
% \begin{equation}
%   \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
% \end{equation}
%
% \begin{equation}
%   \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
% \end{equation}
%
% \begin{equation}
%   \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
% 	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
%   & = v(2-p_{\epsilon}\tau)
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
%   & = \tau p_{\epsilon}
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
% 	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
% 	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
%   & = H - p_{\epsilon}v^{2}\tau
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
% 	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
% 	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
% 	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
% 	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
% \end{split}
% \end{equation}
%
% \begin{equation}
% \begin{split}
%   \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
% \end{split}
% \end{equation}

@@
